---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

```

```{css}

h2 {
  font-weight: bold;
  font-size: 30px;
  text-align: center;
  font-family: "Georgia";
}

h3 {
  font-style: italic;
  font-size: 24px;
  text-decoration: underline;
  font-family: "Georgia";
}

```

## Introduction to the Investigation: Is there an Empty Parking Lot Off Street Every Day of the Week?

This data describes the number of parking lots occupied by parked cars off street from my house from 8am to 11pm from Monday to Sunday. Data of 6 parking lots have been collected, with lots labelled 1, 2, 3, 4, 5 and 6. The data was then analysed to understand which parking lots are occupied by parked cars at which times during each day of the week. Data was collected at 8.00am, 10.00am, 5.00pm, 7.00pm, 9.00pm and 11.00pm each day of the week. These time intervals were chosen to observe if there were any parking lots free at these times as these are the times at which we would need to park out car. With this data, we aim to investigate if my family will be able to park off street our house (out of the 6 parking lots) if we were to get a new car. 

Firstly, the main dataframe that will be used for subsequent data analysis is made from the original data set. The google form data was collected in a google sheet and saved as a csv file. Below, the csv file is read into R. The dataframe is assigned the variable name 'logged_data'. The headers of the table in the dataframe are then renamed. This renamed dataframe is then assigned a new variable name of 'latest_data'. Lastly, the column 'Timestamp' in latest_data is removed, and this new dataset is assigned the variable name 'new_data'. 

```{r, echo=FALSE}

library(tidyverse)
library(magick)
logged_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vS1iF8iHDnxQL_KRHa7ubNPQF84n6OWQ5YHE_wgkQrYl7qBUmFv4slgYZjpp_CCDWfQ0NgCOQdOx_ng/pub?output=csv")
  
latest_data <- logged_data %>% 
  rename(Day = 2,
         Time = 3,
         NumberOfParkingSpotsOccupied = 4,
         ParkingSpotsOccupied = 5)

new_data <- latest_data %>%
  select(-Timestamp)

```

## Tidying of Data

Next, the data in new_data is tidied such that data from each individual day can be analysed. This is done through creating separate data sets for each day of the week (Monday - Sunday). After isolating data collected from each day, the 'Day' column is removed and the 'ParkingSpotsOccupied' column is split into individual rows for each different parking spot chosen in the form. The 'ParkingSpotsOccupied' column is converted to a categorical variable, rather than a continuous one. Finally, any null values in the data set are converted to '0' values. Each different data set is assigned the variable name 'Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday' or 'Sunday'. 

```{r, echo=FALSE}
Monday <- new_data %>% 
  filter(Day == "Monday") %>%
  select(-Day) %>% 
  separate_rows(ParkingSpotsOccupied, sep = ", ") %>%
  mutate(ParkingSpotsOccupied = as.factor(ParkingSpotsOccupied),
         Time = factor(Time, levels = 
                         c("8.00am", "10.00am", "5.00pm", "7.00pm", "9.00pm", "11.00pm")))

Tuesday <- new_data %>%
  filter(Day == "Tuesday") %>%
  select(-Day) %>% 
  separate_rows(ParkingSpotsOccupied, sep = ", ") %>%
  mutate(ParkingSpotsOccupied = replace_na(ParkingSpotsOccupied, "0"),
         ParkingSpotsOccupied = as.factor(ParkingSpotsOccupied),
         Time = factor(Time, levels = 
                         c("8.00am", "10.00am", "5.00pm", "7.00pm", "9.00pm", "11.00pm")))

Wednesday <- new_data %>%
  filter(Day == "Wednesday") %>%
  select(-Day) %>% 
  separate_rows(ParkingSpotsOccupied, sep = ", ") %>%
  mutate(ParkingSpotsOccupied = as.factor(ParkingSpotsOccupied),
         Time = factor(Time, levels = 
                         c("8.00am", "10.00am", "5.00pm", "7.00pm", "9.00pm", "11.00pm")))

Thursday <- new_data %>% 
  filter(Day == "Thursday") %>%
  select(-Day) %>%
  separate_rows(ParkingSpotsOccupied, sep = ", ") %>%
  mutate(ParkingSpotsOccupied = as.factor(ParkingSpotsOccupied),
         Time = factor(Time, levels = 
                         c("8.00am", "10.00am", "5.00pm", "7.00pm", "9.00pm", "11.00pm")))

Friday <- new_data %>% 
  filter(Day == "Friday") %>%
  select(-Day) %>% 
  separate_rows(ParkingSpotsOccupied, sep = ", ") %>%
  mutate(ParkingSpotsOccupied = replace_na(ParkingSpotsOccupied, "0"),
         ParkingSpotsOccupied = as.factor(ParkingSpotsOccupied),
         Time = factor(Time, levels = 
                         c("8.00am", "10.00am", "5.00pm", "7.00pm", "9.00pm", "11.00pm")))

Saturday <- new_data %>%
  filter(Day == "Saturday") %>%
  select(-Day) %>% 
  separate_rows(ParkingSpotsOccupied, sep = ", ") %>%
  mutate(ParkingSpotsOccupied = as.factor(ParkingSpotsOccupied),
         Time = factor(Time, levels = 
                         c("8.00am", "10.00am", "5.00pm", "7.00pm", "9.00pm", "11.00pm")))

Sunday <- new_data %>%
  filter(Day == "Sunday") %>%
  select(-Day) %>% 
  separate_rows(ParkingSpotsOccupied, sep = ", ") %>%
  mutate(ParkingSpotsOccupied = as.factor(ParkingSpotsOccupied),
         Time = factor(Time, levels = 
                         c("8.00am", "10.00am", "5.00pm", "7.00pm", "9.00pm", "11.00pm")))


```

## Plotting Bar Graphs for Each Day of the Week 

Bar graphs were then plotted using the 7 new data sets created. Each bar graph aimed to display the number of parking lots which were occupied by parked cars at each time interval throughout the day. Each bar is filled with the lots which were parked at, distinguished by a different colour as labelled in the key. The times at which the maximum and minimum number of lots were occupied in each day were identified to understand peak and off-peak parking periods. Moreover, the 'total number of cars' parked throughout the day was calculated. Here each lot at each specific time was treated as its own entity. So, even if the same car was parked at the same lot over a few time intervals, it is still assumed to be a different car at a different time interval for the purpose of calculating the frequency each lot was parked at. 

### Monday 

```{r, echo=FALSE}

monday_graph <- Monday %>% 
  ggplot(aes(x = Time, fill = ParkingSpotsOccupied)) + 
  geom_bar(position = "stack") +
  scale_fill_manual(values = c("0" = "#ffa9dc", "1" = "#ffb2a9", "2" = "#ffd4a9", "3" = "#fff7a9", "4" = "#e0ffa9", "5" = "#a9fffc", "6" = "#c8a9ff")) +
  labs(title = "Parking Spots Occupied at from 8.00am - 11.00pm on Monday",
    x = "Time",
    y = "Number of Parking Spots Occupied",
    fill = "Parking Spot Number",
    caption = "Source: Google Forms")
monday_graph

monMax <- new_data %>%
  filter(Day == "Monday") %>%
  filter(NumberOfParkingSpotsOccupied == max(NumberOfParkingSpotsOccupied)) %>%
  select(Time, NumberOfParkingSpotsOccupied)
monMaxTime <- monMax$Time
monMaxSpots <- (monMax$NumberOfParkingSpotsOccupied)[1]

monMin <- new_data %>%
  filter(Day == "Monday") %>%
  filter(NumberOfParkingSpotsOccupied == min(NumberOfParkingSpotsOccupied)) 
monMinTime <- monMin$Time
monMinSpots <- (monMin$NumberOfParkingSpotsOccupied)[1]

monTotalCars <- sum((new_data
                     %>% slice(1:6))$NumberOfParkingSpotsOccupied)
       
```

Parking spot 1 had the highest frequency of parked cars, being occupied throughout the day. Parking spots 3 and 6 had the lowest frequency of parked cars, being vacant throughout the day.

It is observed that the times, `r monMaxTime`, seem to be the most crowded with `r monMaxSpots` parking spots occupied, whereas the time `r monMinTime` seems to be the least crowded with `r monMinSpots` parking spots occupied on Monday. In total, the 6 parking spots were used `r monTotalCars` times. 

### Tuesday 

```{r, echo=FALSE}

tuesday_graph <- Tuesday %>% 
  ggplot(aes(x = Time, fill = ParkingSpotsOccupied)) + 
  geom_bar(position = "stack") + 
  scale_fill_manual(values = c("0" = "#ffa9dc", "1" = "#ffb2a9", "2" = "#ffd4a9", "3" = "#fff7a9", "4" = "#e0ffa9", "5" = "#a9fffc", "6" = "#c8a9ff")) +
  labs(title = "Parking Spots Occupied at from 8.00am - 11.00pm on Tuesday",
       x = "Time",
       y = "Number of Parking Spots Occupied",
       fill = "Parking Spot Number",
       caption = "Source: Google Forms")
tuesday_graph

tueMax <- new_data %>%
  filter(Day == "Tuesday") %>%
  filter(NumberOfParkingSpotsOccupied == max(NumberOfParkingSpotsOccupied)) %>%
  select(Time, NumberOfParkingSpotsOccupied)
tueMaxTime <- tueMax$Time
tueMaxSpots <- (tueMax$NumberOfParkingSpotsOccupied)[1]

tueMin <- new_data %>%
  filter(Day == "Tuesday") %>%
  filter(NumberOfParkingSpotsOccupied == min(NumberOfParkingSpotsOccupied)) %>%
  select(Time, NumberOfParkingSpotsOccupied)
tueMinTime <- tueMin$Time
tueMinSpots <- (tueMin$NumberOfParkingSpotsOccupied)[1]

tueTotalCars <- sum((new_data
                     %>% slice(7:12))$NumberOfParkingSpotsOccupied)

```

Parking spot 1 had the highest frequency of parked cars, being occupied the most throughout the day. Parking spot 6 had the lowest frequency of parked cars, being vancant throughout the day. 

It is observed that the times, `r tueMaxTime`, seem to be the most crowded with `r tueMaxSpots` parking spots occupied, whereas the time `r tueMinTime` seems to be the least crowded with `r tueMinSpots` occupied parking spots on Tuesday. In total, the 6 parking spots were used `r tueTotalCars` times. 


### Wednesday 

```{r, echo=FALSE}

wednesday_graph <- Wednesday %>% 
  ggplot(aes(x = Time, fill = ParkingSpotsOccupied)) + 
  geom_bar(position = "stack") + 
  scale_fill_manual(values = c("0" = "#ffa9dc", "1" = "#ffb2a9", "2" = "#ffd4a9", "3" = "#fff7a9", "4" = "#e0ffa9", "5" = "#a9fffc", "6" = "#c8a9ff")) +
  labs(title = "Parking Spots Occupied at from 8.00am - 11.00pm on Wednesday",
       x = "Time",
       y = "Number of Parking Spots Occupied",
       fill = "Parking Spot Number",
       caption = "Source: Google Forms")
wednesday_graph

wedMax <- new_data %>%
  filter(Day == "Wednesday") %>%
  filter(NumberOfParkingSpotsOccupied == max(NumberOfParkingSpotsOccupied)) %>%
  select(Time, NumberOfParkingSpotsOccupied)
wedMaxTime <- wedMax$Time
wedMaxSpots <- (wedMax$NumberOfParkingSpotsOccupied)[1]

wedMin <- new_data %>%
  filter(Day == "Wednesday") %>%
  filter(NumberOfParkingSpotsOccupied == min(NumberOfParkingSpotsOccupied)) %>%
  select(Time, NumberOfParkingSpotsOccupied)
wedMinTime <- wedMin$Time
wedMinSpots <- (wedMin$NumberOfParkingSpotsOccupied)[1]

wedTotalCars <- sum((new_data
                     %>% slice(13:18))$NumberOfParkingSpotsOccupied)

```

Parking spot 1 had the highest frequency of parked cars, being occupied throughout the day. Parking spot 6 had the lowest frequency of parked cars, only being occupied at 10.00am. All parking spots were used at least once. 

It is observed that the times, `r wedMaxTime`, seem to be the most crowded with `r wedMaxSpots` parking spots occupied, whereas the time `r wedMinTime` seems to be the least crowded with `r wedMinSpots` occupied parking spots on Wednesday. In total, the 6 parking spots were used `r wedTotalCars` times. 

### Thursday

```{r, echo=FALSE}

thursday_graph <- Thursday %>% 
  ggplot(aes(x = Time, fill = ParkingSpotsOccupied)) + 
  geom_bar(position = "stack") + 
  scale_fill_manual(values = c("0" = "#ffa9dc", "1" = "#ffb2a9", "2" = "#ffd4a9", "3" = "#fff7a9", "4" = "#e0ffa9", "5" = "#a9fffc", "6" = "#c8a9ff")) +
  labs(title = "Parking Spots Occupied at from 8.00am - 11.00pm on Thursday",
       x = "Time",
       y = "Number of Parking Spots Occupied",
       fill = "Parking Spot Number",
       caption = "Source: Google Forms")
thursday_graph

thuMax <- new_data %>%
  filter(Day == "Thursday") %>%
  filter(NumberOfParkingSpotsOccupied == max(NumberOfParkingSpotsOccupied)) %>%
  select(Time, NumberOfParkingSpotsOccupied)
thuMaxTime <- thuMax$Time
thuMaxSpots <- (thuMax$NumberOfParkingSpotsOccupied)[1]

thuMin <- new_data %>%
  filter(Day == "Thursday") %>%
  filter(NumberOfParkingSpotsOccupied == min(NumberOfParkingSpotsOccupied)) %>%
  select(Time, NumberOfParkingSpotsOccupied)
thuMinTime <- thuMax$Time
thuMinSpots <- (thuMax$NumberOfParkingSpotsOccupied)[1]

thuTotalCars <- sum((new_data
                     %>% slice(19:24))$NumberOfParkingSpotsOccupied)

```

Parking spots 1 and 2 had the highest frequency of parked cars, being occupied throughout the day. Parking spot 6 had the lowest frequency of parked cars, being vacant throughout the day. 

It is observed that the times, `r thuMaxTime`, seem to be the most crowded with `r thuMaxSpots` parking spots occupied, whereas the time `r thuMinTime` seems to be the least crowded with `r thuMinSpots` parking spots occupied on Thursday. In total, the 6 parking spots were used `r thuTotalCars` times. 


### Friday

```{r, echo=FALSE}

friday_graph <- Friday %>% 
  ggplot(aes(x = Time, fill = ParkingSpotsOccupied)) + 
  geom_bar(position = "stack") + 
  scale_fill_manual(values = c("0" = "#ffa9dc", "1" = "#ffb2a9", "2" = "#ffd4a9", "3" = "#fff7a9", "4" = "#e0ffa9", "5" = "#a9fffc", "6" = "#c8a9ff")) +
  labs(title = "Parking Spots Occupied at from 8.00am - 11.00pm on Friday",
       x = "Time",
       y = "Number of Parking Spots Occupied",
       fill = "Parking Spot Number",
       caption = "Source: Google Forms")
friday_graph

friMax <- new_data %>%
  filter(Day == "Friday") %>%
  filter(NumberOfParkingSpotsOccupied == max(NumberOfParkingSpotsOccupied)) %>%
  select(Time, NumberOfParkingSpotsOccupied)
friMaxTime <- friMax$Time
friMaxSpots <- (friMax$NumberOfParkingSpotsOccupied)[1]


friMin <- new_data %>%
  filter(Day == "Friday") %>%
  filter(NumberOfParkingSpotsOccupied == min(NumberOfParkingSpotsOccupied)) %>%
  select(Time, NumberOfParkingSpotsOccupied)
friMinTime <- thuMin$Time
friMinSpots <- (thuMin$NumberOfParkingSpotsOccupied)[1]

friTotalCars <- sum((new_data
                     %>% slice(25:30))$NumberOfParkingSpotsOccupied)
                    

```

Parking spots 1, 2 and 4 had the highest frequency of parked cars, being occupied throughout the day. Parking spot 6 had the lowest frequency of parked cars, with only 2 cars parked throughout the day. 

It is observed that the times, `r friMaxTime`, seem to be the most crowded with `r friMaxSpots` parking spots occupied, whereas the time `r friMinTime` seems to be the least crowded with `r friMinSpots` parking spots occupied on Friday. In total, the 6 parking spots were used `r friTotalCars` times. 


### Saturday 

```{r, echo=FALSE}

saturday_graph <- Saturday %>% 
  ggplot(aes(x = Time, fill = ParkingSpotsOccupied)) + 
  geom_bar(position = "stack") + 
  scale_fill_manual(values = c("0" = "#ffa9dc", "1" = "#ffb2a9", "2" = "#ffd4a9", "3" = "#fff7a9", "4" = "#e0ffa9", "5" = "#a9fffc", "6" = "#c8a9ff")) +
  labs(title = "Parking Spots Occupied at from 8.00am - 11.00pm on Saturday",
       x = "Time",
       y = "Number of Parking Spots Occupied",
       fill = "Parking Spot Number",
       caption = "Source: Google Forms")
saturday_graph

satMax <- new_data %>%
  filter(Day == "Saturday") %>%
  filter(NumberOfParkingSpotsOccupied == max(NumberOfParkingSpotsOccupied)) %>%
  select(Time, NumberOfParkingSpotsOccupied)
satMaxTime <- satMax$Time
satMaxSpots <- (satMax$NumberOfParkingSpotsOccupied)[1]

satMin <- new_data %>%
  filter(Day == "Saturday") %>%
  filter(NumberOfParkingSpotsOccupied == min(NumberOfParkingSpotsOccupied)) %>%
  select(Time, NumberOfParkingSpotsOccupied)
satMinTime <- satMin$Time
satMinSpots <- (satMin$NumberOfParkingSpotsOccupied)[1]

satTotalCars <- sum((new_data
                     %>% slice(30:35))$NumberOfParkingSpotsOccupied)


```

Parking spots 1 and 2 had the highest frequency of parked cars, being occupied throughout the day. Parking spot 6 had the lowest frequency of parked cars, being vacant throughout the day. 

It is observed that the times, `r satMaxTime`, seem to be the most crowded with `r satMaxSpots` parking spots occupied, whereas the time `r satMinTime` seems to be the least crowded with `r satMinSpots` parking spots occupied on Saturday. In total, the 6 parking spots were used `r satTotalCars` times.  

### Sunday

```{r, echo=FALSE}

sunday_graph <- Sunday %>% 
  ggplot(aes(x = Time, fill = ParkingSpotsOccupied)) + 
  geom_bar(position = "stack") + 
  scale_fill_manual(values = c("0" = "#ffa9dc", "1" = "#ffb2a9", "2" = "#ffd4a9", "3" = "#fff7a9", "4" = "#e0ffa9", "5" = "#a9fffc", "6" = "#c8a9ff")) +
  labs(title = "Parking Spots Occupied at from 8.00am - 11.00pm on Sunday",
       x = "Time",
       y = "Number of Parking Spots Occupied",
       fill = "Parking Spot Number",
       caption = "Source: Google Forms")
sunday_graph

sunMax <- new_data %>%
  filter(Day == "Sunday") %>%
  filter(NumberOfParkingSpotsOccupied == max(NumberOfParkingSpotsOccupied)) %>%
  select(Time, NumberOfParkingSpotsOccupied)
sunMaxTime <- sunMax$Time
sunMaxSpots <- (sunMax$NumberOfParkingSpotsOccupied)[1]


sunMin <- new_data %>%
  filter(Day == "Sunday") %>%
  filter(NumberOfParkingSpotsOccupied == min(NumberOfParkingSpotsOccupied)) %>%
  select(Time, NumberOfParkingSpotsOccupied)
sunMinTime <- sunMin$Time
sunMinSpots <- (sunMin$NumberOfParkingSpotsOccupied)[1]

sunTotalCars <- sum((new_data
                     %>% slice(30:35))$NumberOfParkingSpotsOccupied)


```


Parking spot 1 had the highest frequency of parked cars, being occupied throughout the day. Parking spot 6 had the lowest frequency of parked cars, being vacant throughout the day.  

It is observed that the times, `r sunMaxTime`, seem to be the most crowded with `r sunMaxSpots` parking spots occupied, whereas the time `r sunMinTime` seems to be the least crowded with `r sunMinSpots` parking spots occupied on Sunday. In total, the 6 parking spots were used `r sunTotalCars` times.   


## Evaluation & Conclusion 

It is evident that the most crowded day was Friday, with the parking spots being used a total of 24 times. With parking spot 6 being vacant throughout the day on Monday, Tuesday, Thursday, Saturday and Sunday, we will be able to park our new car in that parking spot 6 on these days. On Wednesday, parking spot 6 remained vacant except at 10.00am. It is possible that spot was taken up by a visiting driver who will not be using that spot frequently. This same assumption can be used for Friday where parking spot 6 was only used at 9.00pm and 11.00pm, given that the spot was vacant at 8.00am on the subsequent Saturday. Moreover, should this assumption not hold, at 10.00am on Wednesday, there were still other parking spots such as 2, 3, 4 and 5 which were vacant and can be used. 

Therefore, I can conclude that should my family choose to get a new car, we will be able to park off street, mainly in parking spot 6.